---
title: "Hwk 4"
author: "Kerry Nixon"
date: "5/22/2021"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(here)
library(tidyverse)
library(dplyr)
library(cowplot)
library(patchwork)
```

As discussed in Homework 3, the Biden Administration recently tasked an Inter-agency Working
Group (IWG) with updating the United States Government's Social Cost of Carbon (SCC). In this
homework, you will use new estimates of the future impacts of climate change to inform an update
to the SCC.

We recommend using R and writing functions to compute your answers wherever possible.
In the provided dataset "damages.csv" you have new model estimates of the annual total damages
from climate change at different levels of warming (in degrees C). The other dataset "warming.csv"
contains estimates of a baseline future climate trajectory (in degrees C) until 2100, and a second
trajectory that adds a one-time pulse of CO2 today to the atmosphere. The pulse is 35 billion tons
of carbon, which is roughly equal to annual global emissions. You can think of this as a "small"
one-time pulse in carbon emissions.

```{r}
#read in the data

dmg <- read.csv(here("data/damages.csv")) %>%
  mutate(warm2 = warming^2) # adding collumn for squared data for quadratic regression

wrm <- read.csv(here("data/warming.csv"))
```


1. Using damages.csv, estimate a quadratic damage function relating the dollar value of damages
to the change in global mean temperature. Omit an intercept term; damages by construction
must equal zero when there is no climate change. Plot your estimated damage function,
overlaid with a scatterplot of the underlying data.


```{r}
data_plot <- ggplot(data = dmg, aes(x = warming, y = damages)) +
  geom_point()

#simple linear model for comparison
dmg_lm <- lm(damages ~ warming, data = dmg)

#dam_lm_quadratic
dmg_lmq <- lm(damages ~ warming + warm2, data=dmg)
dmg_lmq[["coefficients"]][["(Intercept)"]] <- 0


# Compare variance accounted for by each model
summary(dmg_lm)
summary(dmg_lmq)

#plotting the data with dmg_lmq function

data_plot +
  geom_smooth(data = dmg_lm, aes(x = warming, y = damages)) +
  geom_smooth(data = dmg_lmq, aes(x = warming, y = damages), color = "red")

wrm_ceof <- dmg_lmq[["coefficients"]][["warming"]]
wrm2_coef <- dmg_lmq[["coefficients"]][["warm2"]]

```


2. Use warming.csv and your estimated damage function to predict damages in each year under
the baseline climate and the pulse scenario. Make four plots: (1) damages over time without
the pulse, (2) damages over time with the pulse, (3) the difference in damages over time that
arises from the pulse, and (4) the difference in damages over time from the pulse per ton of
CO2 (you can assume that each ton of the pulse causes the same amount of damage).

```{r}
# adding in data
wrm_2 <- wrm %>% 
  mutate(damages_baseline = wrm2_coef*warming_baseline^2 + wrm_ceof*warming_baseline) %>% 
  mutate(damages_pulse = wrm2_coef*warming_pulse^2 + wrm_ceof*warming_pulse) %>% 
  mutate(damage_diff = damages_pulse-damages_baseline) %>% 
  mutate(diff_per_ton = damage_diff/35000000000)

# plotting

# baseline damages
baseline <- ggplot(data = wrm_2, aes(x=year, y=damages_baseline)) +
  geom_line()

# damages pulse
pulse <- ggplot(data = wrm_2, aes(x=year, y=damages_pulse)) +
  geom_line()

# difference in damages
difference <- ggplot(data = wrm_2, aes(x=year, y=damage_diff)) +
  geom_line()

# differenc in damge per tone of CO2
per_ton <- ggplot(data = wrm_2, aes(x=year, y=diff_per_ton)) +
  geom_line()

all_plots <- baseline/pulse | difference/per_ton
all_plots
```


3. The SCC is the present discounted value of the stream of future damages caused by one
additional ton of CO2. The Obama Administration used a discount rate of 3% to discount
damages. Recently, New York State used a discount rate of 2%. Calculate and make a plot of
the SCC (y-axis) against the discount rate (x-axis) for a reasonable range of discount rates.

```{r}
# Calculating SCC for different discount rates

scc <- wrm_2 %>% 
  mutate(pv_1 = diff_per_ton/((1+.01)^X)) %>% 
  mutate(pv_2 = diff_per_ton/((1+.02)^X)) %>% 
  mutate(pv_3 = diff_per_ton/((1+.03)^X)) %>% 
  mutate(pv_4 = diff_per_ton/((1+.04)^X)) %>% 
  mutate(pv_5 = diff_per_ton/((1+.05)^X))

dr <- c(0.01, 0.02, 0.03, 0.04, 0.05)

scc_sum <- data.frame("PV_1" = sum(scc$pv_1),
                      "PV_2" = sum(scc$pv_2),
                      "PV_3" = sum(scc$pv_3),
                      "PV_4" = sum(scc$pv_4),
                      "PV_5" = sum(scc$pv_5)) %>%
  pivot_longer(cols = c(PV_1, PV_2, PV_3, PV_4, PV_5),
               names_to = "dr_percent",
               values_to = "present_damage") %>% 
  replace("dr_percent",dr)

# plot itttttt

ggplot(data = scc_sum, aes(x = dr_percent, y = present_damage)) +
  geom_point(size = 3)

```


4. The National Academies of Sciences, Engineering, and Medicine advised the government in a
2017 report to use the Ramsey Rule when discounting within the SCC calculation:
r = p + ng

Using p = 0.001, n = 2, and g = 0.01, what is the SCC? Locate this point on your graph from
above.

```{r}
# Ramsey Rule with given values
ramsey_r <- 0.001 + 2*0.01

ramsey_r

ramsey <- wrm_2 %>% 
  mutate(pv_r = diff_per_ton/((1+0.021)^X))

dr_ramsey <- c(0.021)

scc_sum_r <- data.frame("PV_r" = sum(ramsey$pv_r)) %>%
  pivot_longer(cols = c(PV_r),
               names_to = "dr_percent",
               values_to = "present_damage") %>% 
  replace("dr_percent",dr_ramsey)

# plot itttttt

ggplot(data = scc_sum, aes(x = dr_percent, y = present_damage)) +
  geom_point(size = 3) +
  geom_point(data = scc_sum_r, color = "red", size = 3)

```


Given the academies' advice for calculating the discount rate, the SCC is \$`r round(scc_sum_r$present_damage,2) ` with a discount rate of `r ramsey_r`



5. Now suppose there are two possible climate policies that can be pursued. Policy A is business as usual and Policy B is to take immediate and strong action on climate change. Use these facts:

• Society is risk neutral

• Use a discount rate of 2%

• If you undertake Policy A there are two possible outcomes. Either warming will occur as in the "baseline" (i.e. "no-pulse") data set above (this happens with probability 0.5) or warming each year will be 1.5 times that in the "baseline" data set (with probability 0.5).

```{r}
# Policy A
wrm_pol_a <- wrm %>% 
  mutate(damages_baseline = wrm2_coef*warming_baseline^2 + wrm_ceof*warming_baseline) %>% 
  mutate(warming_1.5 = (warming_baseline*1.5)) %>% 
  mutate(damages_w1.5 = wrm2_coef*warming_1.5^2 + wrm_ceof*warming_1.5) %>% 
  mutate(scc_baseline = damages_baseline/((1+.02)^X)) %>% 
  mutate(scc_w1.5 = damages_w1.5/((1+.02)^X))

pol_a_basedmg <- sum(wrm_pol_a$scc_baseline)
pol_a_1.5dmg <- sum(wrm_pol_a$scc_w1.5)

pol_a_ev <- (pol_a_basedmg*0.5)+(pol_a_1.5dmg*0.5)
```


• Under Policy B, warming will continue until 2050 as in the "baseline" data set, and then will stabilize at 1.29 degrees and stay that way forever.

```{r}
# Policy B
wrm_pol_b <- wrm %>%  
  mutate(polb_warming = warming_baseline) %>% 
  mutate(polb_warming = case_when(
    year > 2050 ~ 1.29,
    year < 2051 ~ warming_baseline)) %>% 
  mutate(damages_polb = wrm2_coef*polb_warming^2 + wrm_ceof*polb_warming) %>% 
  mutate(scc_polb = damages_polb/((1+0.02)^X))

pol_b_dmg <- sum(wrm_pol_b$scc_polb)

```


What is the expected present value of damages up to 2100 under Policy A? What is the expected present value of damages up to 2100 under Policy B? Suppose undertaking Policy A costs zero and undertaking Policy B costs X. How large could X be for it to still make economic sense to pursue Policy B instead of Policy A? 


```{r}
# maximum cost (X) to make Policy B worth it compared to policy A
max_cost <- pol_a_ev - pol_b_dmg
```


The maximum cost of policy B would be $`r round(max_cost,2)` to continue pursuing policy B. Any additional cost would make policy A more attractive.


Qualitatively, how would your answer change if society were risk averse?

Office hours do you invert the curve for dmages versus gains - looking at the utility (it would be utility lost not utility gained)

